### 2024-06-24 basic-statistics.qmd
20240624 - df

The critical value is the test value for a given test, above which the effect is considered to be [statistically significant](#sec-statisticalSignificance).
The critical value for [statistical significance](#sec-statisticalSignificance) for each test can be determined based on the degrees of freedom and alpha level.

```{r}
alpha <- .05
N <- 200
nGroup1 <- 150
nGroup2 <- 150
numGroups <- 4
numLevelsFactorA <- 3
numLevelsFactorB <- 4
numPredictors <- 5
```

@@

@@

@@

@@

@@

@@

```{r}
df1_RMANOVA <- numGroups - 1
df2_RMANOVA <- (numGroups - 1) * (N - 1)
```

@@

@@

>>>>

The critical value is the test value for a given test, above which the effect is considered to be [statistically significant](#sec-statisticalSignificance).
The critical value for [statistical significance](#sec-statisticalSignificance) for each test can be determined based on the degrees of freedom and alpha level.
The degrees of freedom (*df*) refer to the number of values in the calculation of a test statistic that are free to vary.

```{r}
alpha <- .05
N <- 200
nGroup1 <- 150
nGroup2 <- 150
numGroups <- 4
numLevelsFactorA <- 3
numLevelsFactorB <- 4
numTimepoints <- 4
numPredictors <- 5
```

@@

For a one-sample *t*-test, the degrees of freedom is in @eq-dfOneSampleTtest:

$$
df = N - 1
$$ {#eq-dfOneSampleTtest}

where $N$ is sample size.

@@

For an independent-samples *t*-test, the degrees of freedom is in @eq-dfIndependentSamplesTtest:

$$
df = n_1 + n_2 - 2
$$ {#eq-dfIndependentSamplesTtest}

where $n_1$ is the sample size of group 1 and $n_2$ is the sample size of group 2.

@@

For a paired-samples *t*-test, the degrees of freedom is in @eq-dfPairedSamplesTtest:

$$
df = N - 1
$$ {#eq-dfPairedSamplesTtest}

where $N$ is sample size (i.e., the number of paired observations).

@@

For a one-way ANOVA, the degrees of freedom is in @eq-dfOneWayANOVA:

$$
df_\text{between} = g - 1
df_\text{within} = N - g
$$ {#eq-dfOneWayANOVA}

where $N$ is sample size and $g$ is the number of groups.

@@

For a factorial two-way ANOVA, the degrees of freedom is in @eq-dfFactorialANOVA:

$$
df_\text{Factor A} = a - 1
df_\text{Factor B} = b - 1
df_\text{Interaction} = (a - 1)(b - 1)
df_\text{error} = a * b * (N - 1)
$$ {#eq-dfFactorialANOVA}

where $N$ is sample size, $a$ is the number of levels for factor A, and $b$ is the number of levels for factor B.

@@

For a repeated measures ANOVA, the degrees of freedom is in @eq-dfRepeatedMeasuresANOVA:

$$
df_1 = T - 1
df_2 = (T - 1)(N - 1)
$$ {#eq-dfRepeatedMeasuresANOVA}

where $N$ is sample size and $T$ is the number of timepoints.

```{r}
df1_RMANOVA <- numTimepoints - 1
df2_RMANOVA <- (numTimepoints - 1) * (N - 1)
```

@@

For a correlation, the degrees of freedom is in @eq-dfCorrelation:

$$
df = N - 2
$$ {#eq-dfCorrelation}

where $N$ is sample size.

@@

For multiple regression, the degrees of freedom is in @eq-dfMultipleRegression:

$$
df_1 = p
df_2 = N - p - 1
$$ {#eq-dfMultipleRegression}

where $N$ is sample size and $p$ is the number of predictors.
